---
title: "AAsoildataAnalysis"
author: "conleym"
date: "11/15/2020"
output: html_document
---
```{r setup, include=FALSE}
library(tidyverse)
library(readr)
```

```{r}

AA <- read.csv("~/Chem313Lab01MOC/Lab4Soil/tidy_AA.csv")%>%
mutate(percent_rsd = as.numeric(percent_rsd)*100)
#sample_key <- read.csv("~/Chem313Lab01MOC/Lab4Soil/~Chem313SoilMOC/Sample_Key.csv", skip=0)

```

1) Start by defining useful lists; such as what sites, metals, and sample IDs were analyzed.

```{r}
sample_sites <- unique(filter(AA, site != "MB", site != "")$site)
#excluding method blank and quality control from the list of sites


```

2) Calibrating the AA Data 

```{r Calibration}
#cal_aa <- NULL
#selecting cal standards
cal <- AA %>% 
  filter(type == "CalStd"| type =="CalStd2"| type ==  "CalStd4")%>%
  select(concentration, mean_abs, percent_rsd)

#weighted linear regression
w <- 1/(cal$mean_abs*cal$percent_rsd)^2
model <- lm(cal$mean_abs ~ cal$concentration, weights = w)

slope <- model$coefficients[2]
intercept <- model$coefficients[1]
slope_std <- summary(model)$coefficients[2,2]
intercept_std <- summary(model)$coefficients[1,2]

plot(cal$mean_abs ~ cal$concentration,
       xlab = paste("Concentration of Chromium (ppm)"),
       ylab = "Mean Abs") +
    abline(model, col = "green") +
    title(paste("Calibration for Chromium"))

#i think the code below is where things are going wrong
equation <- tibble(metal = "Cr53", slope, slope_std, intercept, intercept_std)
cal_aa <- rbind(equation)
cal_aa

  
#Clearing the environment (optional, but helps to prevent accidentally using the wrong object!)
#remove(equation, cal, slope, slope_std, intercept, intercept_std, w, model)
```

3) Creating a function to analyze samples
```{r}
sample_analysis <- function(unique_site){
concentration_data <- NULL
#for (unique_sitel in sites_analyzed){
  sample <- filter(AA, site == unique_site)
  data <- NULL
  for(ID in sample$sample_key){
    sample_data <- filter(sample, sample_key == ID)

    #cal <- cal_aa
      
    m <- cal$slope
    b <- cal$intercept
    y <- sample_data$mean_abs
      
    b_e <- cal$intercept_std
    m_e <- cal$slope_std
      
    x <- (y-b)/m
   
    RSD <- ((sample_data$percent_rsd/100)*sample_data$mean_abs)
    ABS <- sample_data$mean_abs
      
    e_yb <- sqrt((RSD)^2 + (b_e)^2)
    yb <- ABS-b
    e_x <- x*sqrt((e_yb/yb)^2 + (m_e/m)^2)
      
    data <- rbind(data, data_frame(sample_key = ID, x, e_x))
      
    if(unique_site != "MB"){
      concentration_data <- data_frame(sample_key = sample_data$sample_key,
                                       analyst = sample_data$analyst,
                                       site = unique_site,
                                       conc_dil = x,
                                       conc_dil_error = e_x) %>%
        rbind(concentration_data)
      }
  #}
  
  if(unique_site == "MB"){
      x <- mean(data$x)
      e_x <- sd(data$x)
      concentration_data <- data_frame(site = unique_site,
                                       conc_dil = x,
                                       conc_dil_error = e_x) %>%
        rbind(concentration_data)
       }
  }
  
  return(concentration_data)
}

```

4) Creating a run_sites function

```{r}
#inputs: a function
#outputs: a data frame with the function outputs from each site
run_sites <- function(Function){
  value <- NULL
  for(site in sample_sites){
    site_value <- Function(site)
    value <- rbind(site_value, value)
  }
  return(value)
}
```

```{r}
MB <- sample_analysis("MB")
uncor_sample <- run_sites(sample_analysis)

MB
uncor_sample
```

```{r}
sample_data_mb <- NULL

  conc_dil_blanked <- uncor_sample$conc_dil-MB$conc_dil
  
  #Error Propagation: subtraction of MB
  conc_dil_blanked_error <- sqrt((uncor_sample$conc_dil_error)^2 + (MB$conc_dil_error))^2
  
  sample_data_mb <- uncor_sample %>% 
    mutate(conc_dil_blanked, conc_dil_blanked_error)%>%
    rbind(sample_data_mb)

sample_data_mb
```

#Sample Prep Procedure

1) mass_soil (example: 1.50621) weighed on analytical balance (uncertainty = ±.001)
*Note: because the analytical lab has two different types of balances, and the uncertainty was not recorded in the data collection we must assume that everyone used the less precise balance*

2) mass_soil was dried in 55˚C oven for 1 week

3) mass_soil was ground with mortar and pestle

4) mass_soil was quantitatively transferred to acid washed teflon beaker (do steps 2-4 affect uncertainty?)

5) mass_soil was digested with ~10 mL MQ water, 3 mL nitric acid, 2 mL hydrochloric acid and heated till steaming for 30 minutes.

*Note: additional 3 mL nitric and 2 mL hydrochloric were added to prevent rxn from running dry*

6) Acid digestion solution was quantitatively transferred to falcon tube and diluted to total_volume which was measured with fill line (example 45 mL) (uncertainty= 1 mL?)

*How could you improve the precision at this step?*

7) Falcon tube centrifuged for 10 minutes at 1500 rpm for 10 min.

8) 10 mL of sample transferred to metals-free 15 mL tube and brought to OHSU for ICPMS analysis. 
(stored in fridge prior to transport)



```{r}
#error propagation
vol_e <- 1
mass_e <- 0.001
dil_1010_e <- sqrt(1^2 + 10^2)
dil_e <- sqrt((dil_1010_e/1010)^2 + (1/10)^2) ## error in 101 dilution factor


sample_data <- merge(AA, sample_data_mb) %>%
  unique()%>%
  mutate(conc_blanked = conc_dil_blanked * (total_volume/1000)/(mass_of_soil/1000),
         conc_blanked_error = conc_blanked*
           sqrt((conc_dil_blanked_error/conc_dil_blanked)^2+
                  (dil_e/101)^2 +
                  (mass_e/mass_of_soil)^2 +
                  (vol_e/total_volume)^2),
         conc_unblanked = conc_dil*(total_volume/1000)/(mass_of_soil/1000),
         conc_unblanked_error = conc_unblanked*
           sqrt((conc_dil_error/conc_dil)^2+
                  (dil_e/101)^2 +
                  (mass_e/mass_of_soil)^2 +
                  (vol_e/total_volume)^2)) %>%
  select(!c(concentration, type, mass_of_soil, total_volume, mean_abs, percent_rsd, conc_dil_blanked, conc_dil_blanked_error, conc_dil, conc_dil_error))

sample_data
```


```{r, warning= FALSE}
rm(list=ls()[!(ls()%in% c("AA", "sample_data"))])

```

#See file 3 for the ICPMS results (statistical analysis and visualization) example code!


Changing code in R & Github (the change you actually want)
  
```

```{r }
warnings()
```
