---
title: "AAsoildataAnalysis"
author: "conleym"
date: "11/15/2020"
output: html_document
---
```{r setup, include=FALSE}
library(tidyverse)
```

```{r}

AA <- read.csv("~/Chem313Lab01MOC/Lab4Soil/tidy_AA.csv")
mutate(percent_rsd = as.numeric(percent_rsd)*100)
```

1) Start by defining useful lists; such as what sites, metals, and sample IDs were analyzed.

```{r}
sample_sites_AA <- unique(filter(AA, site!="MB", site!="")$site)
#excluding method blank and quality control from the list of sites
#Preview the lists to check for potential issues:
sample_sites_AA

```

2) Calibrating the AA Data 

```{r Calibration}
cal <- AA %>%
  filter(type != "Sample", percent_rsd != "HIGH") %>%
  select(mean_abs, percent_rsd, concentration)
#weighted linear regression
w <- 1/(cal$mean_abs*cal$percent_rsd)^2
model <- lm(cal$mean_abs ~ cal$concentration, weights = w)
slope <- model$coefficients[2]
intercept <- model$coefficients[1]
slope_std <- summary(model)$coefficients[2,2]
intercept_std <- summary(model)$coefficients[1,2]
plot(cal$mean_abs ~ cal$concentration,
       xlab = paste("[ ] of Chromium (ppm)"),
       ylab = "Mean Abs") +
    abline(model, col = "green") +
    title(paste("Calibration for Chromium"))
equation <- tibble(metal = "Chromium", slope, slope_std, intercept, intercept_std)
cal <- rbind(equation)
cal
  
AA_cal
#Clearing the environment (optional, but helps to prevent accidentally using the wrong object!)
remove(equation, cal, slope, slope_std, intercept, intercept_std, w, model)
```

3) Creating a function to analyze samples
```{r}
#inputs: unique_site (as a character, ex. "A")
#outputs: concentration vector 
sample_analysis2 <- function(unique_site){
  #unique_site <- "A"
  concentration_data <- NULL
  for (unique_site in sample_sites_AA){
    sample <- filter(AA, site = unique_site)
    data <- NULL 
    
    for(ID in sample$sample_key){ 
      sample_data <- filter(sample, sample_key == ID)
      cal <- filter(AA_cal, site == unique_site) 
  
      #Sample Analysis 
      m <- cal$slope 
      b <- cal$intercept 
      y <- sample_data$cps
  
      b_e<- cal$intercept_std
      m_e <- cal$slope_std
  
      x <- (y-b)/m #The units are dependent on the calibration standards (Kg/mL)
      
      RSD <- sample_data$rsd
      CPS <- sample_data$cps
  
      #Error Propagation
      e_yb <- sqrt((RSD)^2 + (b_e)^2) 
      #error in y-b from the calibration
      yb <- CPS-b
      e_x <- x*sqrt((e_yb/yb)^2+(m_e/m)^2)
      #error in x from the calibration
      
      data <- rbind(data, data_frame(sample_key = ID, x, e_x)) 
      if (unique_site != "MB"){
        concentration_data <- data_frame(sample_key = sample_data$sample_key,
                                         analyst = sample_data$analyst,
                                         metal = unique_metal,
                                         site = unique_site,
                                         conc_dil = x, 
                                         conc_dil_error = e_x)%>%
                              rbind(concentration_data)
      }
    }
    if (unique_site=="MB"){
      x <- mean(data$x)
      e_x <- sd(data$x)
      concentration_data <- data_frame(metal = unique_metal,
                                     site = unique_site, 
                                     conc_dil = x,
                                     conc_dil_error = e_x) %>%
                            rbind(concentration_data)
    }
  }
  return(concentration_data)
}
```

4) Creating a run_sites function
```{r}
#inputs: a function
#outputs: a data frame with the function outputs from each site
run_sites <- function(Function){
  value <- NULL
  for(sites in sample_sites){
    site_value <- Function(sites)
    value <- rbind(site_value, value)
  }
  return(value)
}
```

```{r}
MB2 <- sample_analysis2("MB") #(ug/kg)
uncor_sample2 <- run_sites(sample_analysis2) #values do not account for dilutions (ug/kg)
MB2
uncor_sample
```

```{r}
MB_metal <- MB2$conc_dil

  conc_dil_blanked <- uncor_sample2$conc_dil-mean(MB_metal)
  
  #Error Propagation: subtraction of MB
  conc_dil_blanked_error <- sqrt((uncor_sample2$conc_dil_error)^2 + (sd(MB_metal))^2)
  
  sample_data_mb2 <- uncor_sample2 %>% 
    mutate(conc_dil_blanked, conc_dil_blanked_error)%>%

sample_data_mb2
```

#Sample Prep Procedure

1) mass_soil (example: 1.50621) weighed on analytical balance (uncertainty = ±.001)
*Note: because the analytical lab has two different types of balances, and the uncertainty was not recorded in the data collection we must assume that everyone used the less precise balance*

2) mass_soil was dried in 55˚C oven for 1 week

3) mass_soil was ground with mortar and pestle

4) mass_soil was quantitatively transferred to acid washed teflon beaker (do steps 2-4 affect uncertainty?)

5) mass_soil was digested with ~10 mL MQ water, 3 mL nitric acid, 2 mL hydrochloric acid and heated till steaming for 30 minutes.

*Note: additional 3 mL nitric and 2 mL hydrochloric were added to prevent rxn from running dry*

6) Acid digestion solution was quantitatively transferred to falcon tube and diluted to total_volume which was measured with fill line (example 45 mL) (uncertainty= 1 mL?)

*How could you improve the precision at this step?*

7) Falcon tube centrifuged for 10 minutes at 1500 rpm for 10 min.

8) 10 mL of sample transferred to metals-free 15 mL tube and brought to OHSU for ICPMS analysis. 
(stored in fridge prior to transport)



```{r}
#error propagation
vol_e <- 1
mass_e <- .001

sample_data_AA <- merge(AA,sample_data_mb2)%>% #This adds in important details such as soil mass
  unique()%>%
  mutate(conc_blanked = conc_dil_blanked*(total_volume/1000)/(mass_of_soil/1000),
           conc_blanked_error = conc_blanked* 
           sqrt((conc_dil_blanked_error/conc_dil_blanked)^2+
                  (mass_e/mass_of_soil)^2+ 
                  (vol_e/total_volume)^2),
         conc_unblanked = conc_dil*(total_volume/1000)/(mass_of_soil/1000),
         conc_unblanked_error = conc_unblanked* 
           sqrt((conc_dil_error/conc_dil)^2+
                  (mass_e/mass_of_soil)^2+ 
                  (vol_e/total_volume)^2))%>%
  select(-concentration, #removing unecesssary columns
         -type, 
         -mass_of_soil,
         -total_volume,
         -x.rsd,
         -conc_dil_blanked, 
         -conc_dil_blanked_error, 
         -conc_dil,
         -conc_dil_error)
sample_data_AA
```

```{r}
#group_by(site)
#mean
#95CI
```

```{r, warning= FALSE}
rm(list= ls() [!(ls() %in% c("sampl_date", "sample_data_AA"))])

```

#See file 3 for the ICPMS results (statistical analysis and visualization) example code!


Changing code in R & Github (the change you actually want)
  
```
2.Calculate the average [Pb], [Cr], [Cd], and [As] as measured by ICP-MS from all of the class
data for each soil sample. Be sure to subtract the method blank. Calculate the 95% CI of the
range of concentrations for each metal ( ̅x ± ts /√n).
3. Calculate the average [Pb], [Cr], [Cd], and [As] measured in the class average of method
blank and Buffalo River standard soil quality control check samples. Compare the
concentrations observed in the QC soil sample to reported concentrations (data sheet on the
QC soil is on the moodle).
4. Calculate the average [Cr] as measured by AA from all of the class data for each soil site (A,
B, C, D, E, and F). Be sure to subtract the method blank. Calculate the 95% CI of the range
of concentrations for each metal ( ̅x ± ts /√n).
5. Do the [Cr] measured by AA and ICP-MS agree, within error? To assess this, propagate error
from the uncertainties in each instrument’s calibration curve slope and intercept to obtain a
mean estimate and error.
6. Calculated the average measured concentrations of all metals in each soil sample, in mg / kg
dry soil. This will require accounting for all dilutions and the initial mass of soil digested.
Show your calculation for one of the metals. 

```{r }
